Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-23 src/tex/ms.tex
Working on sec 3

With the loss function evaluated, we apply gradient descent to optimize the tunable
coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
original coefficients given in \citep{khan2016frequency} because they likely lie in the
neighborhood of the minimum that we wish to find. 
We fix our learning rate, $\alpha$, to be $=10^{-6}$, which is small enough to ensure we 
don't move far from the minimum.
Finally, we stop the optimization when the validation loss stops decreasing.
This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.

@@

To evaluate how well the optimization procedure generalize to waveforms that are
not in the training set, we evaluate the mismatch between the fine-tuned model
and the data provided in the SXS catalog for 536 NR waveforms. We select
waveforms that share the same part of the parameter space with the training
waveform, i.e. the waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
and precession (${\chi_{x,y}<5\times10^{-3}}$). Fig.~\ref{fig:intrin_space}
shows how the training and testing waveforms are distributed in the
$q-\chi_{PN}$ space.

To illustrate the effect of optimiziation on an individual waveform level, we
compare the mismatch of a particular waveform before and after optimization with
respect to the NR waveform taken directly from the SXS catalog in Fig.
\ref{fig:0154}. Compared to the original IMRPhenomD waveform, the
optimized waveform has smaller residual from NR waveform both in amplitude and
phase, particularly in the inspiral region, where the amplitude displays a
$50\%$ reduction in error. For a fair comparison, we selected one of the testing
waveforms from the catalog presented in \citep{khan2016frequency}.

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{nl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.

>>>>

With the loss function evaluated, we apply gradient descent to optimize the tunable
coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
original coefficients given in \citep{khan2016frequency} because they likely lie in the
neighborhood of the minimum that we wish to find. 
We fix our learning rate, $\alpha$, to be $=10^{-6}$, which is small enough to ensure we 
don't move far from the minimum.
Finally, we stop the optimization when the validation loss stops decreasing 
(see Sec.~\ref{sec:result} for description of the validation set).
This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.

@@

To evaluate how well the optimization procedure generalizes to waveforms that are
not in the training set, we evaluate the mismatch between the fine-tuned model
and an additional 536 NR waveforms in the SXS catalog i.e., the validation set.
We select waveforms that share the same part of the parameter space with the training
set i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
and precession (${\chi_{x,y}<5\times10^{-3}}$). Figure~\ref{fig:intrin_space}
shows how the training and validation waveforms are distributed in the
$q-\chi_{\rm PN}$ space.

To illustrate the effect of optimiziation on an individual waveform level, in Fig.
\ref{fig:0154} we plot the phase and amplitude of a particular waveform before and after optimization together with
the NR waveform taken directly from the SXS catalog. In the bottom panel one can see that,
compared to the original IMRPhenomD waveform, the optimized waveform has smaller residuals both in amplitude and
phase, particularly in the inspiral region where the amplitude displays a
$50\%$ reduction in error. For a fair comparison, we selected one of the testing
waveforms from the catalog presented in \citep{khan2016frequency}.\te{Which one?}

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Figure~\ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{nl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.
